Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO

### 2016-04-04 Paper/paper.tex
adding text to describe LEA

\newcommand{\chibh}{\chi_\mathrm{BH}}
\newcommand{\chins}{\chi_\mathrm{NS}}
\newcommand{\mbh}{m_\mathrm{BH}}
\newcommand{\mns}{m_\mathrm{NS}}
\newcommand{\mchirp}{\mathcal{M}_c}
\newcommand{\LL}{\mathcal{L}}
\newcommand{\deff}{D_\mathrm{eff}}

@@

\begin{abstract}
We attempt to answer the following questions: 
\begin{enumerate}
 \item What systematic biases are introduced by neglecting the NS tidal
 disruption close to merger? 
 \item At what SNR do the systematic biases dominate over the statistical 
 errors in the recovery of binary parameters?
 \item By using templates which models NS deformation (over inspiral) and 
 NS disruption (close to merger), what information can we recover about
 the $\Lambda_\mathrm{BH}\sim (R/M)^5_\mathrm{NS}$ parameter?
\end{enumerate}
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\vspace{2cm}

\textit{\prayush{Current status:}}

Below we explain the outline the different calculations performed in this 
paper:

\begin{enumerate}
\item We choose the following values of different NSBH binary parameters:
\begin{enumerate}
\item $m_\mathrm{NS}=1.35M_\odot$;
\item $q=m_\mathrm{BH}/\mathrm{NS}=\{2,3\}$;
\item $\chi_\mathrm{NS}=0$;
\item $\chi_\mathrm{BH}=\{-0.5, 0, +0.5\}$;
\item $\Lambda =\{500,1000,2000\}$ ($\Lambda=0\implies \mathrm{BHBH}$).
\end{enumerate}

\item For each system, corresponding to one of all possible unique combinations
of the above parameter choices, we carry out three parameter estimation tests: 
\begin{enumerate}
\item BH-BH signals injected, recovered with BH-BH waveforms (control),
\item NS-BH signals injected, recovered with BH-BH waveforms, (current approach) and
\item NS-BH signals injected, recovered with NS-BH waveforms.
\end{enumerate}
Injections were made at different SNR values: 
\begin{enumerate}
\item $\rho = \{20, 30, 50, 70, 90, 120\}$.
\end{enumerate}
Choice of emcee parameters:
\begin{enumerate}
\item $N_\mathrm{samples}=150,000$; 
\item $N_\mathrm{walkers}=100$; 
\item $N_\mathrm{burn-in}=500$; 
\end{enumerate}
and choice of prior boundaries:
\begin{enumerate} 
\item $m_\mathrm{BH}\in [1.2, 25]M_\odot$;
\item $m_\mathrm{NS}\in [1.2, 15]M_\odot$;
\item $\Lambda\leq 4000$; 
\item $\sigma\left(\Lambda\right) = 100$ (for chains where templates are NS-BH waveforms).
\end{enumerate}

\item Results are shown in
Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}-~\ref{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}.

\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Techniques}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Modeling tidal deformation during inspiral}
Describe the PN terms used here

\subsection{Modeling tidal disruption near merger}
Describe the Lackey et al model

\subsection{Bayesian methods}
% % Describe the emcee-based PE code

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{How are detection searches affected by not including NS matter effects}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%Near future detection searches with aLIGO are poised to search for NSBH
%signals using BHBH templates. In this section, we examine the effect of ignoring
%the effects of tidal deformation and disruption of the NS during inspiral and 
%merger phases of search templates. We construct a template bank $\mathcal{B}$ 
%restricted to the NSBH parameter space with $\mbh\in[2.5M_\odot, 20.5M_\odot]$, 
%$\chibh\in[-0.5, +0.75]$, $\mns\in[1M_\odot,3M_\odot]$ and $\chi_\mathrm{NS}=0$,
%using a PN-based geometric lattice placement method~\cite{Harry:2013tca}. 
%
%
%Before investigating its SNR recovery against tidal NSBH signals, we test the 
%effectualness of this bank to point-particle NSBH signals,
%in order to make sure that the placement method does not leave any region of
%the binary parameter-space under-covered. We simulate $50,000$ signals with 
%component-mass and BH spin values
%sampled uniformly over their allowed ranges, and filter each of them against
%the entire bank. For both signals and templates, we use the frequency-domain 
%reduced-order model (ROM) of the SEOBNRv2 approximant~\cite{Purrer:2014} 
%implemented within the LIGO algorithms Library~\cite{lal}. We record the 
%fitting-factors~\cite{FittingFactorApostolatos} recovered by the bank for all
%signals, which measures the fraction of optimal SNR that the bank is able
%to recover. Fig.~\ref{} shows these FFs as a function of the injected
%BH mass and BH spin, and Fig.~\ref{} provides a histogram of the recovered FF 
%values. From these figures, we conclude that our bank is effectual as intended.
%
%
%Next, tn order to establish the effect of NS matter effects on the SNR recovery of 
%detection searches, we perform a similar calculation as above, with the
%difference that the signals are now modelled using the model of tidal 
%corrections to IMR amplitude and phasing published in 
%Ref.~\cite{Lackey:2013axa}, applied to SEOBNRv2 as the underlying BHBH model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How is PE affected by not including NS matter effects?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{What do we gain by using templates that include NS matter effects?}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Discussion

>>>>

\newcommand{\ii}{\mathrm{i}}
\newcommand{\chibh}{\chi_\mathrm{BH}}
\newcommand{\chins}{\chi_\mathrm{NS}}
\newcommand{\mbh}{m_\mathrm{BH}}
\newcommand{\mns}{m_\mathrm{NS}}
\newcommand{\mchirp}{\mathcal{M}_c}
\newcommand{\LL}{\mathcal{L}}
\newcommand{\deff}{D_\mathrm{eff}}

@@

\begin{abstract}
We attempt to answer the following questions:
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{s1:introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Finally, we note that our results are applicable for the design sensitivity
LIGO instruments, that are expected to come online in $2018-19$. Also note
that more a recent work that improves upon the waveform model we use in this
study has been published~\cite{Pannarale:2015jka}, but it gives only an 
amplitude model and needs to be (in future) augmented with a compatible phase
model. In addition, accuracy of the underlying numerical simulations used to
calibrate the waveform model used here has not been verified against 
independent codes so far.
It is therefore difficult to assess the combined modeling error and the 
results presented in our study are limited by the limitations of the model.
However, we expect the combined effect of modeling errors to not change our
qualitative conclusions.

The remainder of the paper is organized as follows. 
Sec.~\ref{s1:techniques} discusses data analysis techniques and resources 
used in this paper, such as the waveform model, and parameter estimation 
algorithm.
Sec.~\ref{s1:PEwithnoNS} discusses the consequences of ignoring tidal 
effects in parameter estimation waveform models.
Sec.~\ref{s1:PEwithNS} discusses the measurability for the leading order
tidal parameter $\lambdans$ at plausible SNR values.
Sec.~\ref{s1:multiple_observations} discusses the improvement in our
measurement of $\lambdans$ with successive (multiple) observations of
NSBH mergers.
Finally, in Sec.~\ref{s1:discussion} we summarize our results and discuss
future prospects with Advanced LIGO.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Techniques}\label{s1:techniques}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Waveform Models}\label{s2:waveforms}

In Ref.~\cite{Lackey:2013axa}, the authors presented a suite of $134$ numerical
relativity simulations of neutron stars inspiraling into spinning black holes. Amongst
the varied quantities were (a) the equation of state for the NS, (b) NS mass, 
(c) mass-ratio and (d) black hole spins, taken as aligned with the orbit.
A total of $21$ two-parameter EoSs were used for these simulations, which also 
had $q\in[2, 5]$, $\chibh\in[-0.5, +0.75]$, and 
$m_\mathrm{NS}\in[1.20, 1.45]M_\odot$. Based on this suite of simulations, the
authors calibrated a frequency-domain waveform model for the inspiral and merger phasing of
NSBH coalescences. The GW strain predicted by this model (henceforth ``LEA''),
can be written as
% 
\begin{equation}
 \tilde{h}_\mathrm{NSBH}(f, \vec{\theta}) = \tilde{h}_\mathrm{BBH}(f, \vec{\theta} - \{\lambdans\})\,A(f, \vec{\theta})\,e^{\ii \Delta\Phi(f, \vec{\theta})},
\end{equation}
where $\vec{\theta}:=\{M, \eta, \chibh, \lambdans\}$ are the intrinsic binary
parameters. The first term on the right hand side is an underlying
BBH waveform model that is used as a base. In the original construction of LEA,
this was taken to be the SEOBNRv1 model~\cite{Taracchini:2012} of the 
Effective-one-body (EOB) family~\cite{Buonanno99}. The second term changes the 
amplitude of the BBH model
to match that of an NSBH merger of otherwise identical parameters, with NS-matter
effects parametrized by $\lambdans$. During inspiral this term is set to
unity, but is a sensitive function of $\lambdans$ close to merger. The third 
term does the same as the second, but to the waveform phasing. During inspiral,
$\Delta\Phi(f, \vec{\theta})$ is set to the PN tidal phasing corrections,
at the leading and next-to-leading orders~\cite{Vines2011}. Close to merger,
additional phenomenological terms are added and calibrated to the $134$ available
NR simulations.

In this paper we use LEA for our signal and template modeling, but switch the 
underlying BBH model to SEOBNRv2~\cite{Taracchini:2013rva},
using its reduced-order frequency-domain version~\cite{Purrer:2015tud}. We
expect this change to make our conclusions more robust because: (a) the 
SEOBNRv2 model is more accurate~\cite{Kumar:2015tha,Kumar:2016dhh}, and (b)
the differences between the two EOB models are caused by the
inaccuracies of SEOBNRv1 during the {\it inspiral} phase, many orbits before 
merger~\cite{Kumar:2015tha}.
Since LEA only augments the inspiral phasing with PN tidal terms, our
change in the underlying BBH model does not change LEA's construction, but
increases the overall model accuracy during inspiral. Therefore the model used 
here (``LEA+'') is a more robust version of LEA.

\subsection{Bayesian methods}\label{s2:bayesian}
% % Describe the emcee-based PE code

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How is PE affected by not including NS matter effects?}\label{s1:PEwithnoNS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{What do we gain by using templates that include NS matter effects?}\label{s1:PEwithNS}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{s1:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Discussion
